Towards a Neural Co-Processor Which Restores Movement After Stroke: Modeling a Proof-of-Concept

### 2021-11-09 paper/article.tex
Paper progress

\subsubsection{Simulated lesions cause real world failure modes}
Simulating a brain lesion in terms of this artificial network results in error
modes that resemble a natural lesion of certain regions of a primate brain. For example, if we
alter the network by zeroing the outputs of some of the first (input, or AIP) module's neurons,
the reaching motion generally succeeds, but the finger muscle velocities show a high degree
of error - effectively implying that the subject can somewhat reach to grasp, but cannot form a
grasp appropriate for the object. That suggests that losing a portion of the cortical
machinery needed to translate or forward-propagate object shape information
to movement-related cortex can result in a reduced ability to form the hand properly for
the grasp, though positioning of the hand may still roughly succeed. Muscle spasticity of
the hand is also a common symptom of certain strokes in primates, and indeed many who
suffer from it are still able to position their hand, even while being unable to form
it properly for a grasp \cite{khanna.openloop}. In that sense, the error mode of this
simulated lesion closely resembles natural stroke symptoms. Conversely, if we simulate a
``stroke'' of a portion of the output module of the Michaels mRNN, roughly
corresponding to M1, we see a more wholesale loss of movement, affecting even the
ability to reach for the grasp. See Fig. \ref{fig:lesion} for examples.

\begin{figure}
	\centering
	\begin{subfigure}[c]{0.65\textwidth}
	    \centering
	    \includegraphics[width=\textwidth]{lesion_trajs.pdf}
	    \caption{Example muscle trajectories during a single trial, for one hand and shoulder muscle}
	\end{subfigure}
	\hfill
	\begin{subfigure}[c]{0.32\textwidth}
	    \centering
	    \includegraphics[width=\textwidth]{lesion_hand.pdf}
	    \caption{Proportion of loss attributable to hand muscles}
	\end{subfigure}
	\hfill
	\caption{Lesion designs which prevent forward propagation of object shape information
	         differentially impact hand pose. Loss here is a mean squared error (MSE) loss,
		 measured relative to the circuit's trajectories prior to the lesion.}
	\label{fig:lesion}
\end{figure}

The co-processor's task in our simulation will be to identify the appropriate information,
read from the mRNN's activations, for conditioning the stimulation. The co-processor seeks
to effectively bridge across the lesion, forward propatating the object shape information,
indirectly, through the use of stimulation.

In our experiments below, we demonstrate our co-processor design on two types of
simulated lesions:
\begin{itemize}
	\item \textbf{Connection-based:} we prevent the forward and backward propagation of
	      information between the ``F5'' and ``M1'' modules, effectively representing a
	      severing of the connections between the two.
	\item \textbf{Output-based:} we force the output of some proportion of ``M1'' neurons
	      to zero, effectively removing them from the network.
\end{itemize}

TODO: show example trajectories for both

\subsubsection{Simulated network exhibits long running dynamics}
Just as the co-processor must adapt to

\subsubsection{Sensor model}
asdf

\subsubsection{Stimulation model}
asdf

* Overview of test bed
** Michaels model .
** Lesion designs
*** Lesion examples, i.e. hand velocity more affected than shoulder velocity.
*** Disconnect modules vs lesion M1, and the effects
** Stimulation design
*** Spatio-temporal smoothing
** Observation model

* Overview of the four experiments:
** Vanilla
** Co-adaptive with connectivity loss
** Co-adaptive with partial M1 loss
** Co-adaptive with data reuse

\subsection{Training the Model}
Training
** Alternate training EN and CPN
*** Train an EN
**** Training the EN until predictive power reaches a threshold, based on current task performance
**** Training regime based on most recent CPN, similar CPNS (noise added to params), and white noise stimulation
**** Single batch of data from applying current CPN set and white noise to the brain, EN fit to ``offline'' it over maining epochs
**** Predictive power alone is not sufficient for use in training the CPN: we need to form the batch as above, otherwise training is unstable.
*** Train the CPN
**** Backprop through the EN, effectively using the EN's prediction as ground truth

@@

* Any point in showing sensitivity to observation or stim dim?

* Try regularization variation?

>>>>

\subsubsection{Simulated lesions cause real world failure modes}
Simulating a brain lesion in terms of this artificial network results in error
modes that resemble a natural lesion of certain regions of a primate brain. For example, if we
alter the network by zeroing the outputs of some of the first (input, or AIP) module's neurons,
the reaching motion generally succeeds, but the finger muscle velocities show a high degree
of error - effectively implying that the subject can somewhat reach to grasp, but cannot form a
grasp appropriate for the object. That suggests that losing a portion of the cortical
machinery needed to translate or forward-propagate object shape information
to movement-related cortex can result in a reduced ability to form the hand properly for
the grasp, though positioning of the hand may still roughly succeed. Muscle spasticity of
the hand is also a common symptom of certain strokes in primates, and indeed many who
suffer from it are still able to position their hand, even while being unable to form
it properly for a grasp \cite{khanna.openloop}. In that sense, the error mode of this
simulated lesion closely resembles natural stroke symptoms. Conversely, if we simulate a
``stroke'' of a portion of the output module of the Michaels mRNN, roughly
corresponding to M1, we see a more wholesale loss of movement, affecting even the
ability to reach for the grasp. Finally, if we ``disconnect'' communications between the
F5 and M1 modules, we see a failure similar to an AIP lesion: movement is generally
achieved, but we see a disproportionate impact on hand pose.
See Fig. \ref{fig:lesion} for examples.

\begin{figure}
	\centering
	\begin{subfigure}[c]{0.62\textwidth}
	    \centering
	    \includegraphics[width=\textwidth]{lesion_trajs.pdf}
	    \caption{Example muscle trajectories during a single trial, for one hand and shoulder muscle}
	\end{subfigure}
	\hfill
	\begin{subfigure}[c]{0.32\textwidth}
	    \centering
	    \includegraphics[width=\textwidth]{lesion_hand.pdf}
	    \caption{Proportion of loss attributable to hand muscles}
	\end{subfigure}
	\hfill
	\caption{Lesion designs which prevent forward propagation of object shape information
	         differentially impact hand pose. Loss here is an L2 distance
		 measured relative to the circuit's trajectories prior to the lesion.}
	\label{fig:lesion}
\end{figure}

The co-processor's task in our simulation will be to identify the appropriate information,
read from the mRNN's activations, for conditioning the stimulation. The co-processor seeks
to effectively bridge across the lesion, forward propagating the object shape information,
indirectly, through the use of stimulation.

In our experiments below, we demonstrate our co-processor design on three types of
simulated lesions:
\begin{itemize}
	\item \textbf{Connection-based:} we prevent the forward and backward propagation of
	      information between the ``F5'' and ``M1'' modules, effectively representing a
	      severing of the connections between the two.
	\item \textbf{Output-based (AIP):} we force the output of some proportion of ``AIP''
	      neurons to zero, effectively removing them from the network. This results in a
	      loss of object shape information and, as we will see below, causes the co-processor
	      to be unable to differentiate objects.
	\item \textbf{Output-based (M1):} we force the output of some proportion of ``M1''
	      neurons to zero, causing a more holisitic loss of movement.
\end{itemize}

\subsubsection{Simulated network exhibits long running dynamics}
Just as the co-processor must adapt to the lesion, it must also adapt to the network
dynamics.

\begin{figure}
	\centering
	\begin{subfigure}[c]{0.48\textwidth}
	    \centering
	    \includegraphics[width=\textwidth]{perturbe_single.pdf}
	    \caption{}
	\end{subfigure}
	\hfill
	\begin{subfigure}[c]{0.48\textwidth}
	    \centering
	    \includegraphics[width=\textwidth]{perturbe_dist.pdf}
	    \caption{}
	\end{subfigure}
	\hfill
	\caption{Instantaneous perturbations of a random sample (n=10) of M1 neurons at time t=200 results in long-running effects on output.}
	\label{fig:dynamics}
\end{figure}

\subsubsection{Observation model}
** Observation model
*** Spatio-temporal smoothing

\begin{figure}
\includegraphics[width=\textwidth]{stim_and_obs.png}
\caption{Example of stimulation and observation for a single trial}
\centering
\label{fig:stim_and_obs}
\end{figure}

\subsubsection{Stimulation model}
** Stimulation design

\subsection{Experiments}
Summary of tunables:
\begin{itemize}
	\item Learning rate schedules
	\item Activation functions
	\item Obs function dimensionality
	\item Stim dimensionality
	\item Dimensionality of EN, CPN
\end{itemize}

Experiment types:
\begin{itemize}
	\item Connectivity lesion
	\item Co-adaptive with connectivity lesion
	\item Co-adaptive with partial M1 loss
	\item Co-adaptive with partial AIP loss
	\item Co-adaptive with data reuse
\end{itemize}

\begin{figure}
\includegraphics[width=\textwidth]{cpn_michaels_arch_labeled.eps}
\caption{Experiment architecture overview}
\centering
\label{fig:exp_overview}
\end{figure}

\subsection{Training the Model}
Alternate training EN and CPN.
\begin{itemize}
	\item Train an EN
		\begin{itemize}
			\item Training the EN until predictive power reaches a threshold, based on current task performance
			\item Training regime based on most recent CPN, similar CPNS (noise added to params), and white noise stimulation
			\item Single batch of data from applying current CPN set and white noise to the brain, EN fit to ``offline'' it over maining epochs
			\item Predictive power alone is not sufficient for use in training the CPN: we need to form the batch as above, otherwise training is unstable.
		\end{itemize}
	\item Train the CPN
		\begin{itemize}
			\item Backprop through the EN, effectively using the EN's prediction as ground truth
		\end{itemize}
\end{itemize}

@@
